Numerical trajectory patterns of floating macroalgae

### 2020-09-23 thesis_chapter_3.Rmd
added morphology graphs and summary stats for area estimation

@@

The use of shape coefficients was used as a novel approach to simulating biological drifting debris. The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical. The cross-sectional area for each of the virtual kelp parts was determined and used in the calculation of hydrodynamic drag. The hydrodynamic/form drag can be calculated by the equation: $$ F_d = 0.5 \rho A_c C_d U^2$$

@@

p_particle <- tidync("data/trajectory_data/PassiveFloat.nc") %>%
  hyper_tibble() %>%
  dplyr::select(lon, lat, time) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))

@@

k_particle <- tidync("data/trajectory_data/KelpFloat.nc") %>%
  hyper_tibble() %>%
  dplyr::select(lon, lat, time) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))

>>>>

## Cross-sectional area

In order to incorporate hydrodynamic and wind drag, the cross-sectional area of the kelp was calculated first. Known geometric shapes reflecting the relevant plant sections were used to estimate the surface area for various parts of the plant, for details please see image ???. The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin] and estimated from figure \ref{fig:morphology}.

```{r morphology_wrangling, message=FALSE, warning=FALSE, include=FALSE}
# Investigate morphology data to assist in estimating surface area
# Load software
library(tidyverse)
library(ggpubr)
library(gridExtra)
library(ggplot2)
library(ggord)
library(vegan)
# Read in morphology data
eck_data <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
  filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
  filter(site!= "De Hoop")
eck_data$site <- as.character(eck_data$site)
eck_deep <- eck_data %>% 
  filter(population == "Deep")

eck_long <- gather(eck_deep, morphology, measurement, frond_mass:total_length, factor_key=TRUE)
# Summary stats to make it easier to refer to later, these will not be used any analysis.
eck_stats <- eck_long %>% 
  group_by(site, morphology) %>% 
  summarise(mean = mean(measurement, na.rm = T),
            min = min(measurement, na.rm = T),
            max = max(measurement, na.rm = T),
            range = range(measurement, na.rm = T)[2]-range(measurement, na.rm = T)[1],
            sd = sd(measurement, na.rm = T)) %>% 
  ungroup()
# Make custom positions
positionsEck <- c("Betty's Bay", "Kalk Bay", "A-Frame", 
                  "Miller's Point", "Baboon Rock", "Batsata Rock", 
                  "Buffelsbaai", "Scarborough", 
                  "Soetwater","Kommetjie","Hout Bay", 
                  "Oudekraal", "St Helena Bay")
# Make plots for each morphology feature and grouped by site.
frond_mass <- ggplot(data = eck_data, aes(x = site, y = frond_mass, fill = site)) +
           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
           stat_compare_means(method = "kruskal.test", label.x = 5) + 
           scale_x_discrete(limits = positionsEck) + 
           ylab("Frond mass (g)") + 
           coord_flip() +
           xlab("") + 
           theme_bw()
primary_length <- ggplot(data = eck_data, aes(x = site, y = primary_length , fill = site)) +
           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
           stat_compare_means(method = "kruskal.test", label.x = 5) + 
           scale_x_discrete(limits = positionsEck) + 
           ylab("Primary Length  (cm)") + 
           coord_flip() +
           xlab("") + 
           theme_bw()
primary_width <- ggplot(data = eck_data, aes(x = site, y = primary_width , fill = site)) +
           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
           stat_compare_means(method = "kruskal.test", label.x = 5) + 
           scale_x_discrete(limits = positionsEck) + 
           ylab("Primary Width (cm)") + 
           coord_flip() +
           xlab("") + 
           theme_bw() 
frond_length <- ggplot(data = eck_data, aes(x = site, y = frond_length , fill = site)) +
           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
           stat_compare_means(method = "kruskal.test", label.x = 5) + 
           scale_x_discrete(limits = positionsEck) + 
           ylab("Frond length (cm)") + 
           coord_flip() +
           xlab("") + 
           theme_bw() 
stipe_mass <- ggplot(data = eck_data, aes(x = site, y = stipe_mass , fill = site)) +
           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
           stat_compare_means(method = "kruskal.test", label.x = 5) + 
           scale_x_discrete(limits = positionsEck) + 
           ylab("Stipe mass (g)") + 
           coord_flip() +
           xlab("") + 
           theme_bw() 
stipe_length <- ggplot(data = eck_data, aes(x = site, y = stipe_length , fill = site)) +
           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
           stat_compare_means(method = "kruskal.test", label.x = 5) + 
           scale_x_discrete(limits = positionsEck) + 
           ylab("Stipe length (cm)") + 
           coord_flip() +
           xlab("") + 
           theme_bw() 
stipe_circ <- ggplot(data = eck_data, aes(x = site, y = stipe_circ , fill = site)) +
           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
           stat_compare_means(method = "kruskal.test", label.x = 5) + 
           scale_x_discrete(limits = positionsEck) + 
           ylab("Stipe circ (cm)") + 
           coord_flip() +
           xlab("") + 
           theme_bw() 
tufts <- ggplot(data = eck_data, aes(x = site, y = tufts , fill = site)) +
           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
           stat_compare_means(method = "kruskal.test", label.x = 5) + 
           scale_x_discrete(limits = positionsEck) + 
           ylab("No. of tufts") + 
           coord_flip() +
           xlab("") + 
           theme_bw() 
epi_length <- ggplot(data = eck_data, aes(x = site, y = epi_length , fill = site)) +
           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
           stat_compare_means(method = "kruskal.test", label.x = 5) + 
           scale_x_discrete(limits = positionsEck) + 
           ylab("Epi Length  (cm)") + 
           coord_flip() +
           xlab("") + 
           theme_bw()

total_length <- ggplot(data = eck_data, aes(x = site, y = epi_length, fill = site)) +
           geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE, show.legend = FALSE) +
           stat_compare_means(method = "kruskal.test", label.x = 5) + 
           scale_x_discrete(limits = positionsEck) + 
           ylab("Total length  (cm)") + 
           coord_flip() +
           xlab("") + 
           theme_bw()
```

```{r morphology, echo=FALSE, , fig.height=7, fig.width=8, message=FALSE, warning=FALSE, fig.cap='\\label{fig:morphology}Boxplots representing *E. pallida* morphological characteristics grouped by site around the Western Cape coastline, with the *y*-axis depicting the specific morphology measured and with units provided. The lower and upper "hinges" correspond to the first and third quartiles. The whiskers represent the range, solid black lines represent the median and black dots are outliers.'}
# Plot the morphology data based on site (see previous section)
ggarrange(frond_length, frond_mass,primary_length, primary_width, stipe_circ, stipe_length, stipe_mass, total_length, tufts, common.legend = TRUE, legend = c("none"), align = c("hv"))
```

@@

The use of shape coefficients was used as a novel approach to simulating biological drifting debris. The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical. The cross-sectional area for each of the virtual kelp parts was determined and used in the calculation of hydrodynamic drag. The hydrodynamic/form drag can be calculated by: $$ F_d = 0.5 \rho A_c C_d U^2$$

@@

p_particle <- tidync("OceanParcels/trajectory_data/PassiveFloat.nc") %>%
  hyper_tibble() %>%
  dplyr::select(lon, lat, time) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))

@@

k_particle <- tidync("OceanParcels/trajectory_data/KelpFloat.nc") %>%
  hyper_tibble() %>%
  dplyr::select(lon, lat, time) %>%
  dplyr::rename(t = time) %>%
  mutate(t = as.POSIXct(t, origin = "2002-01-01 00:00:00"))
